TEMPERATURES AND METALLICITIES OF M DWARFS IN THE \apogee\ SURVEY

### 2019-07-31 paper/draft.tex
intro grammar

\author[0000-0002-6523-9536]{Adam Burgasser}
\affil{Center for Astrophysics and Space Science, University of California San Diego, La Jolla, CA 92093, USA}

@@

Low-mass stars (with masses $M_{*} < 0.7\,\msun$ and atmospheric temperatures $T_{\rm eff} < 4000\,$K) are by far the most ubiquitous type of star, comprising $\sim 70 \%$ of the galaxy's population by number \citep{Bochanski:2010}. With nuclear fusion timescales of $\tau>10^{11}$ yr \citep{Laughlin:1997}, chemical compositions of M dwarf populations tracing the nucleosynthetic processes and interstellar mixing of heavy elements from many generations of shorter-lived, high mass stars, are a unique probe for piecing together galactic structure and evolution \citep{Bochanski:2010,Woolf:2012}.

Additionally, the low masses of M dwarfs make for easier detection of planets by radial velocity offset detections \citep{Trifonov:2018}, high planet-to-star radii ratios make for easier detection of exoplanet transits in light curve observations \citep{Nutzman:2008}, and shorter orbital periods (for a fixed stellar insulation flux) allow for new planet discovery in less observation time than their higher mass counterparts. For these reasons, M dwarfs are primary candidates for exoplanet searches, including by the NASA \kepler\ \citep[e.g.,][]{Dressing2015} and Transiting Exoplanet Survey Satellite (\tess) \citep[e.g.,][]{Muirhead:2018} missions. As a result, detailed and precise knowledge of M dwarf chemical compositions has become key to constraining the the properties, formation scenarios, and atmospheric conditions of potentially habitable exoplanets observable with {\sl JWST}.

Advances in instrumentation with the launch of several spectroscopic surveys in the past decade, such as the Sloan Digital Sky Survey (SDSS), and the Large Sky Area Multi-Object Fiber Spectroscopic Telescope (LAMOST) have dramatically increased the sample of known M dwarfs (\citealt{West:2011}, \citealt{Guo:2015} with catalogs of over 70,000 sources), enabling studies of M dwarf abundances on both exoplanet and Galactic scales.
The launch of the Apache Point Observatory Galactic Evolution Experiment (\apogee; \citealt{Majewski:2015}) survey as part of the Sloan Digital Sky Survey III/IV mission \citep{Eisenstein:2011,Blanton:2017} has introduced the largest sample of M dwarfs observed with high resolution spectroscopy \citep{Desphande:2013}. \apogee\ pipeline measurements of $\teff$ and $\feh$ \citep{Perez:2016} for M dwarfs have only been determined to 100\,K and 0.18\,dex down to $T_{\rm eff}\sim3550\,$K using atmospheric models \citep{Schmidt:2016}.

Elemental abundance measurements from high resolution spectra of F, G, and K stars have achieved extremely high precision (down to $0.01-0.03$\,dex; \citealt{Nissen2018}) enabled by improvements in atmospheric models, including realistic 3D local thermodynamic equilibrium (LTE) assumptions \citep{Asplund2005}, and differential abundance techniques using equivalent widths \citep{Bedell2014}. However, the determination of precise metallicities for M dwarfs has remained a long-standing challenge due to the formation of complex diatomic and triatomic molecular bands at M dwarf temperature regimes, particularly TiO and VO dominating the optical spectrum, H$_2$O and CO for the infrared, and hydride bands (FeH, CaH, CrH, MgH, etc.) present in the spectra of the latest spectral types \citep{Allard1997}.
In particular, atmospheric models fail to reproduce spectral features \citep[e.g.,][]{Mann:2013c} due to incomplete line lists and opacities, and the presence of millions of weak, blended transitions, and the lack of continuum in a spectrum, all make it difficult to deconvolve individual features and extract line strengths from equivalent widths. The combination of these effects creates fundamental limitations to explore the information content available out of high resolution spectra using a traditional comparison to atmosphere models.

A number of studies focused on improving M dwarf metallicity precisions have used systems of M dwarfs in common proper motion with an FGK star, to developed precise empirical relations (as good as $\sim$0.07\,dex) between the adopted metallicities of the FGK primary (under the assumption kinematically associated wide binaries are co-evolutionary; \citealt{Bonfils:2005}), and strong, isolated lines in the spectra of the M dwarf (e.g. \citealt{Terrien:2012}, \citealt{Rojas-Ayala:2012}, \citealt{Newton:2014}). However these metallicity calibrations are inaccessible for analysis with \apogee, as the wavelength coverage does not include important calibration indices in the $K$-band and does not contain information about the overall spectral shape often used to determine $\teff$ and spectral type. Further, earlier calibrations are generally based on moderate-resolution data \citep[with some exceptions,][]{Neves:2014, Lindgren:2016} that fail to utilize the additional spectral information provided by \apogee's resolution.

>>>>

\author[0000-0002-6523-9536]{Adam J. Burgasser}
\affil{Center for Astrophysics and Space Science, University of California San Diego, La Jolla, CA 92093, USA}

@@

Low-mass stars, with masses $M_{*} < 0.7\,\msun$ and effective temperatures $T_{\rm eff} < 4000\,$K, are by far the most ubiquitous type of star, comprising $\sim 70 \%$ of the Galaxy's population by number \citep{Bochanski:2010}. With nuclear fusion timescales $\tau>10^{11}$ yr \citep{Laughlin:1997}, the chemical compositions of the M dwarf populations trace the nucleosynthetic processes and interstellar mixing of heavy elements from many generations of shorter-lived, high mass stars, and are a unique probe for piecing together galactic structure and evolution \citep{Bochanski:2010,Woolf:2012}.

Additionally, the low masses of M dwarfs make for easier detection of planets by radial velocity variability \citep{Trifonov:2018}, high planet-to-star radii ratios make for easier detection of exoplanet transits in light curve observations \citep{Nutzman:2008}, and shorter orbital periods (for a fixed stellar insolation flux) allow for new planet discovery in less observation time than for higher mass stars. For these reasons, M dwarfs are primary candidates for exoplanet searches, including by the NASA \kepler\ \citep[e.g.,][]{Dressing2015} and Transiting Exoplanet Survey Satellite (\tess) \citep[e.g.,][]{Muirhead:2018} missions. As a result, detailed and precise knowledge of M dwarf chemical compositions has become key to constraining the the properties, formation scenarios, and atmospheric conditions of potentially habitable exoplanets observable with the James Webb Space Telescope ({\sl JWST;} \citealt{Clampin2008}).

Advances in instrumentation with the launch of several spectroscopic surveys in the past decade, such as the Sloan Digital Sky Survey \citep[SDSS;][]{Eisenstein:2011,Blanton:2017}, and the Large Sky Area Multi-Object Fiber Spectroscopic Telescope \citep[LAMOST;][]{Zhao:2012} have dramatically increased the sample of known M dwarfs \citep{West:2011,Guo:2015} with catalogs of over 70,000 sources, enabling studies of M dwarf abundances on a Galactic scale.
The Apache Point Observatory Galactic Evolution Experiment (\apogee; \citealt{Majewski:2015}) survey as part of the SDSS III/IV mission has introduced the largest sample of M dwarfs observed with high resolution spectroscopy \citep{Desphande:2013}. \apogee\ pipeline measurements of $\teff$ and $\feh$ \citep{Perez:2016} for M dwarfs have been determined to precisions 100\,K and 0.18\,dex down to $T_{\rm eff}\sim3550\,$K using atmospheric models \citep{Schmidt:2016}.

Elemental abundance measurements from high resolution spectra of F, G, and K stars have achieved extremely high precision (down to $0.01-0.03$\,dex; \citealt{Nissen2018}) enabled by improvements in atmospheric models, including realistic 3D local thermodynamic equilibrium (LTE) assumptions \citep{Asplund2005}, and differential abundance techniques using equivalent widths \citep{Bedell2014}. However, the determination of precise metallicities for M dwarfs has remained a long-standing challenge due to the formation of complex diatomic and triatomic molecular bands at M dwarf temperatures, particularly TiO and VO bands dominating the optical spectrum, H$_2$O and CO bands in the infrared, and hydride bands (FeH, CaH, CrH, MgH, etc.) present in the spectra of the latest spectral types \citep{Allard1997}.
Atmospheric models often fail to reproduce these spectral features \citep[e.g.,][]{Mann:2013c} due to incomplete line lists and opacities, and the presence of millions of weak, blended transitions, and the absence of a clear continuum, all contribute to make it difficult to deconvolve individual features and extract line strengths from equivalent widths. The combination of these effects limit our ability to explore the information content of high resolution spectra using traditional comparison to atmosphere models.

A number of studies focused on improving M dwarf metallicity precisions have used systems of M dwarfs in common proper motion with an FGK star and strong, isolated lines in the spectra of the M dwarf (e.g. \citealt{Terrien:2012}, \citealt{Rojas-Ayala:2012}, \citealt{Newton:2014}) to developed precise empirical relations (as good as $\sim$0.07\,dex). However these metallicity calibrations are inaccessible for analysis with \apogee, as the wavelength coverage does not include important calibration indices in the $K$-band nor information about the overall spectral shape often used to determine $\teff$ and spectral type. Further, earlier calibrations are generally based on moderate-resolution data \citep[with some exceptions,][]{Neves:2014, Lindgren:2016} that fail to utilize the greater spectral information provided by \apogee's resolution.
